What can GitHub and Googlecode Issues tell us about Exceptions in Android and Java?

### 2014-08-31 android-stacks.tex
changes all over

\author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1},
Georgios Gousios\IEEEauthorrefmark{2}, Arie van Deursen\IEEEauthorrefmark{2},
Lucas Almeida\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Federal University of Rio Grande do Norte\\
Natal, Brazil\\
Email: \texttt{roberta@dimap.ufrn.br,lucas.almeida@ppgsc.ufrn.br}}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\
Delft, The Netherlands\\
Email: \texttt{\{g.gousios,arie.vandeursen\}@tudelft.nl}}
}

@@

This paper reports a repository mining study which analyzed all the issues created on
Android projects available in GitHub (482 applications) and Google code (157 applications)
Overall, our research set includes 32,582 issues in GitHub and 41,674 issues in Googlecode from which (4,208)  (1,986) in GC stack traces 
were extracted. Such stack traces were distilled and analyzed in 
combination with source code and bytecode analysis. This study 
consistently identifies common characteristics of reported 
exception stacks traces - e.g., the NullPointerException was the reported cause 
on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
50\% of all projects analyzed. Moreover,  the cross-project analysis 
of exception stack traces enabled to consistently reveal   
fragilities of exception handling code: 
(1) undocumented runtime exceptions signaled by third party code,
(2) undocumented checked exceptions thrown by Android native code
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
Such fragilities impose challenges to application robustness not only 
in Android environment but also to other systems based on Java exception model.

@@

Along the recent years, we have witnessed an astonishing increase in the number 
mobile applications being developed and some of them being largely used. 
Such increase is motivated by pervasiveness of the World Wide Web 
and the unprecedented increase of global smartphone shipments -
the mobile industry shipped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.

@@

In the same rate the number of a users increase, also increases 
the number of users affected by application faults and crashes.
If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face new threats to application robustness imposed by mobile environment.
Such applications have to cope with an increasing number of abnormal
computation states, arising from: programming errors;
 noisy user inputs; faults in underlying middleware or hardware; 
faults on third party services and libraries, compatibility issues [ref], 
connection with devices e.g., as GPS, camera, memory and battery restrictions
among others.

@@

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps.
% as any other modern system~\cite{bruntink2006discovering}. 
The Android platform uses the exception handling mechanism embedded
 in Java to deal with such conditions. Although the exception handling mechanism
~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
As a consequence that the exception handling code may become itself a source of faults, 
and may affect the system the other way around, making the system even less robust ~\cite{robillard2003static,coelho2011unveiling}.

@@

Our conjecture is that several Android applications may be suffering from similar 
fragilities on the exception handling code (of the applications themselves or 
the underlaying platform), and that such fragilities may be leading to
decrease the application robustness, leading for instance to the Uncaught Exception problem
 are one of the main causes of application crashes ~\cite{jo2004uncaught} ~\cite{zhang2012amplifying}.

@@

The goal of this study was to extract exception stack traces embedded on issues of Android projects 
(hosted in GitHub and GoogleCode), looking for its 
common characteristics and investigating whether
 they could reveal common fragilities on the exception code of both the 
applications and the underlaying  framework.

@@

  \item[RQ2] \noindent\emph{Which patterns of exception (mis)-use emerge?}
Based on a set of best practices compiled in this study, the study investigates whether characteristics on exception stack traces  
can pinpoint best practices violations. Characteristics of exception stack traces and detected violations can reveal fragilities
 on the exception handling code, which threats the application dependability.

@@

@@

\noindent\emph{Repository Connectors}  This component is responsible to connect to
 the host repositories in order to collect project issues. The GitHub Connector connected
 to GiHub through GHTorrent. The GoogleCode connector comprised of a Web Crawler,
 that traversed Googlecode Web interface in order to extract the project issues.

\noindent\emph{Exception Stack Trace Distiller} 
Distills the information that composes a stack trace and storing the results in a
relational database. Some of the information extracted from the stack traces were:
 the exception being thrown, its signaler, the exception root cause, its corresponding signaler,
and packages related to each one of these fields.
The tool is based on a combination between a regular expression based parser 
with heuristics able to identify and filter exception names and stack traces inline with text. In
contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can extract all causes related to an exception on a stack,
and stack traces embedded in logs files.
AS PRESENTED IN FIGURE X

@@

\noindent\emph{Exception Signaler Classifier} 
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Plaform, Android Libcore, Aplication, Library). 
We downloaded the source code of Android Platform and OS (level 19)
and this module extracts the packages hierarchies (recursive list of 
of Java packages), that composes it. To discover the packages that compose 
each application, this module extracted the manifest files of each Android apps
 - which defines the main packages that compose the applications.
Then based on pattern matching between the signaler name and the packages 
 this module identifies the origin of the exception signalers.
The exceptions were considered from libraries if their packages were not defined 
on Android Platform, core libraries nor on applications.
Table ~ref{signalers} summarizes the signaler classification.

@@

\subsubsection{Manual Inspections}
This study was also based in a set of manual inspection tasks. Besides the 
 identification of Android projects details previously. 
Manual inspections were used discover the type of exceptions thrown by 
the applications and libraries we manually inspect the sourcecode or javadoc
 documentation of each of them. When the exception was removed from the 
last app version, the previous versions were analyzed. When the exception was
 not found on the previous versions, the exception was discarded.  Only 3 exceptions 
remained undefined and were removed from this study.

@@

@@

 However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
violating the declaration. The Java compiler, could not detect this violation because does 
not perform static exception  checking on native methods. This type of bugs is hard to diagnose
because the developer usualy do not have access to the native code (written in other language).

Consequently, since it was not expected by the programmer, when such method throws 
this exception, such undocumented exception may remain
uncaught and cause the app to crash, or be mistakenly handled by subsumption - 
The exception stack traces reporting this scenario is actually a real bug of Android 
Gingerbread version (which still accounts for 13.6% of devices running Android).

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
  0,8\% of the methods documented the explicitly signaled runtime exceptions.
  The study also detected an undocumented checked exception signaled by
  a native method defined in Android.
}
\end{minipage}}

@@

{\footnotesize
\begin{verbatim}

1. private Activity performLaunchActivity(...) {
2.  try {
3.   ...
4.   ...callActivityOnCreate(...);
5.  } catch (Exception e) {
6.    ...
7.   throw new RuntimeException(..., e);
8.  }
9.}
\end{verbatim}
}

@@

{\footnotesize
\begin{verbatim}
protected void done() {
 ...
   try {
      ...
   } catch (InterruptedException e) {
      android.util.Log.w(..., e);
   } catch (ExecutionException e) {
      throw new RuntimeException("...",e.getCause());
   } catch (CancellationException e) {
      ...
   } catch (Throwable t) {
     throw new RuntimeException("...", t);
   }
    ...
 }
}
\end{verbatim}
}

@@

 %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
\begin{tabular}{ll}
    \hline
 \bfseries{Runtime-Error Wrapping Examples} &   \\  %\bfseries{\#}
    \hline
java.lang.RuntimeException - java.lang.OutOfMemoryError & \\ %30
java.lang.RuntimeException - java.lang.NoClassDefFoundError &\\ %6
java.lang.RuntimeException - java.lang.UnsatisfiedLinkError) & \\ %5
java.lang.RuntimeException -  java.lang.StackOverflowError)  & \\ %4
\hline
\bfseries{Checked-Runtime Wrapping Examples} &  \\
 \hline
...android.robospice.CacheSavingException - java.lang.ClassCastException &  \\ %6
java.lang.reflect.InvocationTargetException - java.lang.NullPointerException &   \\ %3
java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException &  \\ %2
java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException &  \\ %2
 \hline
\bfseries{Checked-Error Wrapping Examples} & \\
 \hline
java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError &  \\ %5
java.lang.Exception - java.lang.OutOfMemoryError &  \\ %1
java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError &	 \\ %1
java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError&	 \\ %1
\hline
\bfseries{Error-Checked Wrapping Examples} & \\
 \hline
java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException &  \\ %4
java.lang.AssertionError - javax.crypto.ShortBufferException) &  \\ %3
\hline
\bfseries{Error-Runtime Wrapping Examples} &   \\
 \hline
java.lang.ExceptionInInitializerError - java.lang.NullPointerException &  \\ %9
java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException &	 \\ %2
 \hline
  \end{tabular}
\caption{Examples of Cross-type wrappings}
\label{tab:wrapping01}
\end{table}

@@

\begin{figure} \centering \includegraphics[scale=0.4]{codeexample2.pdf}
\caption{Code snippets illustrating fragilities on the exception handling code} \label{fig:exchier} \end{figure}

@@

Hence, when (mis)applied, the exception wrapping can make the exception handling code
 more complex and lead to what we call an \emph{exception handling confusion problem}.
Such problem can lead the program to
an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
 one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the JVM which make impossible the program to recover.

@@

\textit{Tools Support for Detecting Faults on Java Exception Handling Code.}

Manually analyzing the exception handling code looking for defects, can easily 
become infeasible [ref]. Hence, some tools have been proposed to detect faults 
on the exception handling code either statically (cabral, roberta,nelio,chang) 
or dynamicaly through testing (roberta). Both approaches, however, have inherent
 limitations. On the one hand, the static analysis tools needs to deal with the false 
positives and cannot adequately deal with dynamically loaded code (e.g. reflection  
[ref]). On the other hand the testing tools are llimited to the fact that the developer
 needs to create a test to exercise every exception, which is also costly.... (olhar tese). 
Moreover, both methods fail to account for exceptions implicitly signaled by the runtime 
environment (e.g. due to resource depletion which can be signaled in every statement invocation).

>>>>

\author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1},
Georgios Gousios\IEEEauthorrefmark{2}, 
Arie van Deursen\IEEEauthorrefmark{2},
Lucas Almeida\IEEEauthorrefmark{1}
}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Federal University of Rio Grande do Norte\\
Natal, Brazil\\
Email: \texttt{roberta@dimap.ufrn.br,lucas.almeida@ppgsc.ufrn.br}}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\
Delft, The Netherlands\\
Email: \texttt{\{g.gousios,arie.vandeursen\}@tudelft.nl}}
}

@@

This paper reports a repository mining study which analyzed all the issues created on
Android projects available in GitHub (482 applications) and Google code (157 applications)
Overall, our research set includes 32,582 issues in GitHub and 41,674 issues in Googlecode from which (4,208)  (1,986) in GC stack traces 
were extracted. Such stack traces were distilled and analyzed in 
combination with source code and bytecode analysis. This study 
consistently identifies common characteristics of reported 
exception stacks traces - e.g., the NullPointerException was the reported cause 
on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
50\% of all projects analyzed. Moreover,  the cross-project analysis 
of exception stack traces enabled to consistently reveal   
fragilities of the exception-related code: 
(1) undocumented runtime exceptions signaled by third party code,
(2) undocumented checked exceptions thrown by Android native code
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
Such fragilities impose challenges to application robustness not only 
in Android environment but also to other systems based on Java exception model.

@@

Along the recent years, we have witnessed an astonishing increase in the number 
mobile applications being developed and some of them becoming largely used. 
Such increase is motivated by pervasiveness of the World Wide Web 
and the unprecedented increase of global smartphone shipments -
the mobile industry shipped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.

@@

In the same rate the number of a users increase, also increases 
the number of users affected by application faults and crashes.
If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face threats to application robustness.
Such applications have to cope with an increasing number of abnormal
computation states, arising from:
 noisy user inputs; faults in underlying middleware or hardware; 
faults on third party services and libraries; compatibility issues [ref]; 
connection with devices e.g., as GPS, camera; memory and battery restrictions; 
 programming errors, among others.

@@

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps.
% as any other modern system~\cite{bruntink2006discovering}. 
The Android platform uses the exception handling mechanism embedded
 in Java to deal with such conditions. Although the exception handling mechanism
~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
As a consequence that the exception-related code may become itself a source of faults, 
and may affect the system the other way around, making the system even less robust ~\cite{robillard2003static,coelho2011unveiling}.

@@

Our conjecture is that several Android applications may be suffering from similar 
fragilities on the exception-related code (of the applications themselves or 
the underlaying platform), and that such fragilities may be leading to
decrease the application robustness, leading for instance to the Uncaught Exception problem -
 which is one of the main causes of application crashes ~\cite{jo2004uncaught} ~\cite{zhang2012amplifying}.

@@

The goal of this mining study was to extract exception stack traces embedded on issues of Android projects 
(hosted in GitHub and GoogleCode), looking for its 
common characteristics and investigating whether
 they could reveal common fragilities on the exception code of both the 
applications and the underlaying  framework.

@@

%Which patterns of exception (mis)use emerge?
  \item[RQ2] \noindent\emph{Which fragilities on the exception-related code emerge from them?}
Based on a set of best practices compiled in this study, we investigated whether characteristics of exception stack traces  
can pinpoint best practices violations. Such  violations reveal fragilities
on the exception-related code, which threats the application dependability.

@@

%Characteristics of exception stack traces and detected violations can reveal fragilities
% on the exception-related code, which threats the application dependability.

@@

\noindent\emph{Repository Connectors.}  This component is responsible to connect to
 the host repositories in order to collect project issues. The GitHub Connector connected
 to GiHub through GHTorrent. The GoogleCode connector comprised of a Web Crawler,
 that traversed Googlecode Web interface in order to extract the project issues.

\noindent\emph{Exception Stack Trace Distiller.} 
Distills the information that composes a stack trace and storing the results in a
relational database. Some of the information extracted from the stack traces were:
 the exception being thrown, its signaler, the exception root cause, its corresponding signaler,
and packages related to each one of these fields.
The tool is based on a combination between a regular expression based parser 
with heuristics able to identify and filter exception names and stack traces inline with text. In
contrast to existing issue parsing solutions such as Infozilla, the parser
created in this work can extract all causes related to an exception on a stack,
and stack traces embedded in logs files.
AS PRESENTED IN FIGURE X

@@

\noindent\emph{Exception Signaler Classifier.} 
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Plaform, Android Libcore, Aplication, Library). 
We downloaded the source code of Android Platform and OS (level 19)
and this module extracts the packages hierarchies (recursive list of 
of Java packages), that composes it. To discover the packages that compose 
each application, this module extracted the manifest files of each Android apps
 - which defines the main packages that compose the applications.
Then based on pattern matching between the signaler name and the packages 
 this module identifies the origin of the exception signalers.
The exceptions were considered from libraries if their packages were not defined 
on Android Platform, core libraries nor on applications.
Table ~ref{signalers} summarizes the signaler classification.

@@

\subsection{Manual Inspections}
This study was also based in a set of manual inspection tasks. Besides the 
 identification of Android projects details previously. 
Manual inspections were used discover the type of exceptions thrown by 
the applications and libraries we manually inspect the sourcecode or javadoc
 documentation of each of them. When the exception was removed from the 
last app version, the previous versions were analyzed. When the exception was
 not found on the previous versions, the exception was discarded.  Only 3 exceptions 
remained undefined and were removed from this study.

@@

%\begin{figure*}
%\centering
%\includegraphics[width=\hsize]{overview.pdf}
%\caption{Overview of the study based on mixed approaches.}
%\label{fig:overviewfig}
%\end{figure*}

@@

 However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
violating the declaration. The Java compiler, could not detect this violation because it does 
not perform static exception checking on native methods. This type of bug is hard to diagnose
because the developer usualy do not have access to the native implementations.

Consequently, since it is not expected by the programmer, when such method throws 
this exception, such undocumented exception may remain
uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
The exception stack traces reporting this scenario is actually a real bug of Android 
Gingerbread version (which still accounts for 13.6% of devices running Android).

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
 Only 0,8\% of the methods documented the explicitly signaled runtime exceptions.
  The study also detected an undocumented checked exception signaled by
  a native method defined in Android.
}
\end{minipage}}

@@

%MOVED TO THE PICTURE
%{\footnotesize
%\begin{verbatim}
%1. private Activity performLaunchActivity(...) {
%2.  try {
%3.   ...
%4.   ...callActivityOnCreate(...);
%5.  } catch (Exception e) {
%6.    ...
%7.   throw new RuntimeException(..., e);
%8.  }
%9.}
%\end{verbatim}
%}

@@

%{\footnotesize
%\begin{verbatim}
%protected void done() {
% ...
 %  try {
 %     ...
  % } catch (InterruptedException e) {
  %    android.util.Log.w(..., e);
  % } catch (ExecutionException e) {
  %    throw new RuntimeException("...",e.getCause());
  % } catch (CancellationException e) {
  %    ...
 %  } catch (Throwable t) {
 %    throw new RuntimeException("...", t);
 %  }
 %   ...
% }
%}
%\end{verbatim}
%}

@@

 %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
\begin{tabular}{ll}
    \hline
 id & \bfseries{Runtime-Error Wrapping Examples}    \\  %\bfseries{\#}
    \hline
1 & java.lang.RuntimeException - java.lang.OutOfMemoryError  \\ %30
2 &java.lang.RuntimeException - java.lang.NoClassDefFoundError \\ %6
3 &java.lang.RuntimeException - java.lang.UnsatisfiedLinkError)  \\ %5
4& java.lang.RuntimeException -  java.lang.StackOverflowError)   \\ %4
\hline
 & \bfseries{Checked-Runtime Wrapping Examples}   \\
 \hline
5& ...android.robospice.CacheSavingException - java.lang.ClassCastException   \\ %6
6&java.lang.reflect.InvocationTargetException - java.lang.NullPointerException    \\ %3
7&java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException  \\ %2
8&java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException   \\ %2
 \hline
& \bfseries{Checked-Error Wrapping Examples}  \\
 \hline
9&java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError  \\ %5
10&java.lang.Exception - java.lang.OutOfMemoryError   \\ %1
11&java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError 	 \\ %1
12&java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError	 \\ %1
\hline
& \bfseries{Error-Checked Wrapping Examples}  \\
 \hline
13&java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException   \\ %4
14&java.lang.AssertionError - javax.crypto.ShortBufferException)   \\ %3
\hline
& \bfseries{Error-Runtime Wrapping Examples}    \\
 \hline
15&java.lang.ExceptionInInitializerError - java.lang.NullPointerException   \\ %9
16&java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException 	 \\ %2
 \hline
  \end{tabular}
\caption{Examples of Cross-type wrappings}
\label{tab:wrapping01}
\end{table}

@@

\begin{figure*} \centering \includegraphics[scale=0.55]{codeexamples.pdf}
\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:exchier} \end{figure*}

@@

Hence, when (mis)applied, the exception wrapping can make the exception-related code
 more complex and lead to what we call an \emph{exception handling confusion problem}.
Such problem can lead the program to
an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
 one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the JVM which make impossible the program to recover.

@@

\textit{Tools Support for Detecting Faults on Java Exception-Related Code.}

Manually analyzing the exception-related code and looking for defects, can easily 
become infeasible ~\cite{Robil00}. Hence, some tools have been proposed to detect faults 
on the exception handling code either statically (cabral, roberta,nelio,chang) 
or dynamicaly through testing (roberta). Both approaches, however, have inherent
 limitations. On the one hand, the static analysis tools needs to deal with the false 
positives and cannot adequately deal with dynamically loaded code (e.g. reflection  
[ref]). On the other hand the testing tools are llimited to the fact that the developer
 needs to create a test to exercise every exception, which is also costly.... (olhar tese). 
Moreover, both methods fail to account for exceptions implicitly signaled by the runtime 
environment (e.g. due to resource depletion which can be signaled in every statement invocation).
